---
title: "september 2023 Bikeshare, Thomas Robbins"
output: html_notebook
---

Loading necessary packages from R Studio library:
```{r}

library(tidyverse)
library(janitor)
library(here)
library(rwunderground)
library(openmeteo)

```

Downloading necessary raw data from 'data_raw':
```{r}

thisfile=here("data_raw", "202309-capitalbikeshare-tripdata.zip")

df1=read_csv(thisfile) %>% clean_names()

```

Using scatter plot to visualize the locations of bikeshare trip starts:
```{r}

df1 |>
  ggplot(aes(x = start_lng, y = start_lat), color = rideable_type) + 
  labs(x = "Trip Start Longitude", y = "Trip Start Latitude") +
  geom_point(aes(alpha = 0.1), size = 2)

```
Creating df2 for analysis of ridership over time:
```{r}

df2 = df1 |> 
  pivot_longer(
    cols = c(started_at, ended_at),
    names_to = "start_end",
    values_to = "t") |> 
  arrange(t) |> 
  mutate(rider_delta=(start_end == "started_at")*2-1) |> 
  mutate(riders=cumsum(rider_delta)) |> 
  relocate(t,start_end,riders, .after = rideable_type)

```


Ridership over time in September 2023:
```{r}

df2 |> 
  ggplot(aes(t,riders)) +
  geom_line() +
  ggtitle("Riders in September")

```

Ridership over time for a single example date (Sept 18):
```{r}

df2 |> 
  filter(day(t)==18) |> 
  ggplot(aes(t,riders)) +
  geom_line() +
  ggtitle("Riders on Sept 18th")

```


```{r}

df2 |> 
  filter(month(t)==9) |> 
  ggplot(aes(t,riders)) +
  geom_line() +
  facet_wrap(~mday(t), scales = "free_x", ncol = 7) +
  ggtitle("Riders by day-of-month for September 2023")

```


Looking at differences between weekdays:
```{r}

df2 |> 
  filter(month(t)==9) |> 
  mutate(epi=epiweek(t),wd=wday(t,label=TRUE,week_start = 7)) |> 
  ggplot(aes(hms::as_hms(t),riders,color = as.factor(epi))) +
  geom_line(alpha = 0.7) +
  facet_wrap(~wd, scales = "free_x", ncol = 7) +
  ggtitle("Ridership over time, by weekday, for Sept 2023")

```


Electric bikes vs classic bikes:
```{r}

df3 = df2 |> 
  mutate(type_riders = cumsum(rider_delta), .by = rideable_type) |> 
  relocate(type_riders, .after = riders)

df3 |> 
  filter(month(t) == 9) |> 
  ggplot(aes(t, type_riders, color = rideable_type)) +
  geom_line() + facet_wrap(~mday(t), scales = "free_x", ncol = 7)

```


Pulling September weather data from R package library:
```{r}

#weather_forecast("Washington",
#hourly = c("temperature_2m","precipitation"),
#response_units = list(
#  temperature_unit = "fahrenheit",
#  precipitation_unit = "inch"))

df_4 = weather_history("Washington",
start = "2023-09-01",
end = "2023-09-30",
hourly = c("apparent_temperature",
           "wind_speed_10m",
           "precipitation",
           "rain"))

```











